---
title: "EPL Datat Analysis"
author: "Jack Kenyon Brown"
format: html
editor: visual
embed-resources: true
---

# EPL Data Analysis

As a part of "The Impact of Video Assistant Referee (VAR) on the English Premier League" by Jack Kenyon Brown, et. al.

```{r}
#| label: Load Packages
#| echo: false
#| warning: false

library(tidyverse)
library(lme4)
library(MASS)
library(stats)
library(car)
library(readxl)
library(writexl)
library(gt)
library(cowplot)
library(dplyr)
library(multcompView)
library(asht)
library(usethis)
library(nortest)
library(gridExtra)
```

## Data Prep

```{r}
#| label: Import, Format, and Filter Data
#| output: false
#| echo: true

#Import EPL_Totals (make sure TotalxG and totalcorners are imported as numeric)


#Be sure to assign Season, Week, and PostVAR variables as factors
EPL_Totals$Season <- factor(EPL_Totals$Season, levels = unique(EPL_Totals$Season))
EPL_Totals$Wk <- factor(EPL_Totals$Wk, levels = unique(EPL_Totals$Wk))
EPL_Totals$PostVAR <- factor(EPL_Totals$PostVAR, levels = unique(EPL_Totals$PostVAR))


#1 year comparison data
EPL_1 <- EPL_Totals %>%
  filter(Season %in% c("18/19", "19/20"))

#3 year comparison data
EPL_3 <- EPL_Totals %>%
  filter(Season %in% c("16/17", "17/18", "18/19", "19/20", "20/21", "21/22"))

#5 year comparison data
EPL_5 <- EPL_Totals
```

## Variables of Interest

|                              |                 |
|:----------------------------:|:---------------:|
|        ***Variable***        | ***Code Name*** |
|            Goals             |   totalgoals    |
|            Fouls             |  totalfoulscom  |
|    Penalty Kicks Awarded     |   totalpksatt   |
|     Penalty Kicks Scored     |  totalpksmade   |
| Penalty Kick Conversion Rate | totalpkconvrate |
|            Shots             |   totalshots    |
|           Offsides           |    totaloff     |
|          Red Cards           |    totalred     |
|         Yellow Cards         |   totalyellow   |
|           Tackles            | totaltackleswon |
|        Interceptions         |    totalint     |

## -

## Checking Normality Assumption

```{r}
#|: Shapiro-Wilk, Kolmogorov-Smirnov, and Anderson-Darling Tests for Normality
#| warning: false

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalpkconvrate", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create an empty dataframe to store the results
test_results <- data.frame(Variable = character(), Shapiro_Wilk_p = numeric(), KS_p = numeric(), AD_p = numeric(), stringsAsFactors = FALSE)

# Loop over each variable
for (var in variables) {
  # Extract the data for the current variable
  data <- na.omit(EPL_Totals[[var]])
  
  # Perform Shapiro-Wilk test
  shapiro_result <- shapiro.test(data)
  
  # Perform Kolmogorov-Smirnov test
  ks_result <- ks.test(data, "pnorm", mean = mean(data), sd = sd(data))
  
  # Perform Anderson-Darling test
  ad_result <- ad.test(data)
  
  # Append the results to the dataframe
  test_results <- rbind(test_results, data.frame(
    Variable = var,
    Shapiro_Wilk_p = shapiro_result$p.value,
    KS_p = ks_result$p.value,
    AD_p = ad_result$p.value
  ))
}
```

### Histograms of Response Variables

```{r}
# Specify the variables
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", 
               "totalpkconvrate", "totalshots", "totaloff",  
               "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create a list to store the plots
plots <- list()

# Generate a histogram for each variable and store it in the list
for (var in variables) {
  p <- ggplot(EPL_Totals, aes_string(x = var)) + 
    geom_histogram(binwidth = 1, fill = "blue", color = "black") +
    theme_minimal()
  plots[[var]] <- p
}

# Arrange the plots in a 4x3 grid
do.call("grid.arrange", c(plots, ncol = 3))
```

## 

## -

## Z-Tests

```{r}
z_test <- function(sample_data, post_var, population_sd) {
  # Split data into two groups based on PostVAR
  group1 <- sample_data[post_var == 0]
  group2 <- sample_data[post_var == 1]
  
  # Calculate means and sample sizes
  mean1 <- mean(group1, na.rm = TRUE)
  mean2 <- mean(group2, na.rm = TRUE)
  n1 <- length(group1)
  n2 <- length(group2)
  
  # Calculate z-statistic
  z_statistic <- (mean2 - mean1) / sqrt((population_sd^2 / n1) + (population_sd^2 / n2))
  
  # Calculate p-value
  p_value <- 2 * (1 - pnorm(abs(z_statistic)))
  
  # Confidence interval (assuming a normal distribution)
  conf_int <- (mean2 - mean1) + c(-1, 1) * qnorm(0.975) * sqrt((population_sd^2 / n1) + (population_sd^2 / n2))
  
  list(estimate = mean2 - mean1, conf.int = conf_int, p.value = p_value)
}

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalpkconvrate", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Assume a population standard deviation for each variable (you need to set this based on your knowledge of the data)
population_sd <- c(
  totalgoals = 1.6434029, totalfoulscom = 5.1740743, totalpksatt = 0.5122013, totalpksmade = 0.4553075, totalpkconvrate = 0.3898591, 
  totalshots = 5.7816171, totaloff = 2.2169412, totalcross = 9.0693778, totalred = 0.3562214, totalyellow = 2.0222872, 
  totaltackleswon = 7.9990457, totalint = 8.3870906)

# Create an empty list to store z-test results
z_tests <- list()

# Loop over each variable
for (var in variables) {
  # 1 year
  z_test_1 <- z_test(EPL_1[[var]], EPL_1$PostVAR, population_sd[[var]])
  # 3 years
  z_test_3 <- z_test(EPL_3[[var]], EPL_3$PostVAR, population_sd[[var]])
  # 5 years
  z_test_5 <- z_test(EPL_5[[var]], EPL_5$PostVAR, population_sd[[var]])
  
  # Extract z-test estimate, 95% CI, and p-value, and store in data frame
  result_df <- data.frame(
    Variable = var,
    Statistic = c("Estimate", "Lower CI limit", "Upper CI limit", "P-value"),
    `1 year` = c(z_test_1$estimate, z_test_1$conf.int[1], z_test_1$conf.int[2], z_test_1$p.value),
    `3 years` = c(z_test_3$estimate, z_test_3$conf.int[1], z_test_3$conf.int[2], z_test_3$p.value),
    `5 years` = c(z_test_5$estimate, z_test_5$conf.int[1], z_test_5$conf.int[2], z_test_5$p.value)
  )
  
  # Append the result dataframe to the list
  z_tests[[var]] <- result_df
}

# Combine all result data frames into one
z_test_results <- do.call(rbind, z_tests)
```

## 

## T-Tests

```{r}
#|: lable: T-Test Analysis

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalpkconvrate", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create an empty list to store t-test results
t_test_results <- list()

# Loop over each variable
for (var in variables) {
  # 1 year
  t_test_1 <- t.test(get(var) ~ PostVAR, data = EPL_1)
  # 3 years
  t_test_3 <- t.test(get(var) ~ PostVAR, data = EPL_3)
  # 5 years
  t_test_5 <- t.test(get(var) ~ PostVAR, data = EPL_5)
  
  # Extract t-test estimate, 95% CI, and p-value, and store in data frame
  result_df <- data.frame(
    Variable = var,
    `1 year` = c(t_test_1$estimate, t_test_1$conf.int, t_test_1$p.value),
    `3 years` = c(t_test_3$estimate, t_test_3$conf.int, t_test_3$p.value),
    `5 years` = c(t_test_5$estimate, t_test_5$conf.int, t_test_5$p.value)
  )
  
  #Store results in a list
  t_test_results[[var]] <- result_df
}

# Write t-test results to an Excel file with different tabs
write_xlsx(t_test_results, "/Users/jackbrown/Desktop/Thesis/T Test Results.xlsx")

```

## Mann-Whitney U-Tests

```{r}
#|: lable: Mann-Whitney U-Test Analysis

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalpkconvrate", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create an empty dataframe to store all t-test results
MW_U_results <- data.frame()

# Loop over each variable
for (var in variables) {
  # 1 year
  t_test_1 <- wmwTest(get(var) ~ PostVAR, data = EPL_1)
  # 3 years
  t_test_3 <- wmwTest(get(var) ~ PostVAR, data = EPL_3)
  # 5 years
  t_test_5 <- wmwTest(get(var) ~ PostVAR, data = EPL_5)
  
  # Extract relevant information and store in data frame
  result_df <- data.frame(
    Variable = var,
    Test = c("Estimate", "Conf.Lower", "Conf.Upper", "P.Value"),
    `1_year` = c(t_test_1$estimate, t_test_1$conf.int[1], t_test_1$conf.int[2], t_test_1$p.value),
    `3_years` = c(t_test_3$estimate, t_test_3$conf.int[1], t_test_3$conf.int[2], t_test_3$p.value),
    `5_years` = c(t_test_5$estimate, t_test_5$conf.int[1], t_test_5$conf.int[2], t_test_5$p.value)
  )
  
  # Append results to the master dataframe
  MW_U_results <- bind_rows(MW_U_results, result_df[4,])
}
```

## Generalized Linear Models

```{r}
#| label: Generalized Linear Models for all three datasets/time period comparisons
#| warning: false

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalpkconvrate", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create an empty dataframe to store all t-test results
GLM_results <- data.frame()

# Loop over each variable
for (var in variables) {
  # 1 year
  glm_1 <- glm.nb(get(var) ~ PostVAR, link = "log", data = EPL_1)
  # 3 years
  glm_3 <- glm.nb(get(var) ~ PostVAR, link = "log", data = EPL_3)
  # 5 years
  glm_5 <- glm.nb(get(var) ~ PostVAR, link = "log", data = EPL_5)
  
  # Extract relevant information and store in data frame
  result_df <- data.frame(
    Variable = var,
    Test = c("p-value", "AIC"),
    `1_year` = c(summary(glm_1)$coefficients[8], summary(glm_1)$aic),
    `3_years` = c(summary(glm_3)$coefficients[8], summary(glm_3)$aic),
    `5_years` = c(summary(glm_5)$coefficients[8], summary(glm_5)$aic)
  )
  
  # Append results to the master dataframe
  GLM_results <- bind_rows(GLM_results, result_df[1:2,])
}

write_xlsx(GLM_results, "/Users/jackbrown/Desktop/Thesis/GLM Results.xlsx")
```

## -

## Appendix

#### Seperate Analysis for Corner Kicks

```{r}
EPL_2 <- EPL_Totals %>%
  filter(Season %in% c("17/18", "18/19", "19/20", "20/21"))
```

**T-Tests**

```{r}
t.test(totalcorners ~ PostVAR, data = EPL_2)
```

**Mann-Whitney U Tests**

```{r}
wmwTest(totalcorners ~ PostVAR, data = EPL_2)
```

**Generalized Linear Models**

```{r}
summary(glm.nb(totalcorners ~ PostVAR, link = "log", data = EPL_2))$coefficients[8]
```
