---
title: "EPL Datat Analysis"
author: "Jack Kenyon Brown"
format: html
editor: visual
embed-resources: true
---

# EPL Data Analysis

As a part of "The Impact of Video Assistant Referee (VAR) on the English Premier League" by Jack Kenyon Brown, et. al.

```{r}
#| label: Load Packages
#| echo: false
#| warning: false

library(tidyverse)
library(readxl)
library(writexl)
library(stats)
library(nortest)
library(gridExtra)
library(car)
library(lmtest)
library(broom)
library(dplyr)
library(asht)
library(MASS)
library(usethis)
```

## Data Prep

```{r}
#| label: Import, Format, and Filter Data
#| output: false
#| echo: true

#This analysis requires the "EPL_Totals" dataset, the can be obtained using the web scraping tools found in the "FBref EPL Web Scraper.qmd" found in the GitHub repository


#Import EPL_Totals dataset


#Be sure to assign Season, Week, and PostVAR variables as factors
EPL_Totals$Season <- factor(EPL_Totals$Season, levels = unique(EPL_Totals$Season))
EPL_Totals$Wk <- factor(EPL_Totals$Wk, levels = unique(EPL_Totals$Wk))
EPL_Totals$PostVAR <- factor(EPL_Totals$PostVAR, levels = unique(EPL_Totals$PostVAR))
#Setting totalcorners as numeric
EPL_Totals$totalcorners <- as.numeric(EPL_Totals$totalcorners)

#1 year comparison data
EPL_1 <- EPL_Totals %>%
  filter(Season %in% c("18/19", "19/20"))

#3 year comparison data
EPL_3 <- EPL_Totals %>%
  filter(Season %in% c("16/17", "17/18", "18/19", "19/20", "20/21", "21/22"))

#5 year comparison data
EPL_5 <- EPL_Totals
```

## Variables of Interest

|                       |                 |
|:---------------------:|:---------------:|
|    ***Variable***     | ***Code Name*** |
|         Goals         |   totalgoals    |
|         Fouls         |  totalfoulscom  |
| Penalty Kicks Awarded |   totalpksatt   |
| Penalty Kicks Scored  |  totalpksmade   |
|         Shots         |   totalshots    |
|       Offsides        |    totaloff     |
|       Red Cards       |    totalred     |
|     Yellow Cards      |   totalyellow   |
|        Tackles        | totaltackleswon |
|     Interceptions     |    totalint     |

## -

## Checking Normality Assumption

### SW, AD, and KS Tests

```{r}
#| label: Shapiro-Wilk, Anderson-Darling, and Kolmogorov-Smirnov Tests for Normality
#| warning: false

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create an empty dataframe to store the results
normality_results <- data.frame(Dataset = character(), Variable = character(), Shapiro_Wilk_p = numeric(), KS_p = numeric(), AD_p = numeric(), stringsAsFactors = FALSE)

# List of datasets
datasets <- list(EPL_1 = EPL_1, EPL_3 = EPL_3, EPL_5 = EPL_5)

# Loop over each variable and dataset
for (dataset_name in names(datasets)) {
  for (var in variables) {
    # Extract the data for the current variable from the current dataset
    data <- na.omit(datasets[[dataset_name]][[var]])
    
    # Perform Shapiro-Wilk test
    shapiro_result <- shapiro.test(data)
    
    # Perform Kolmogorov-Smirnov test
    ks_result <- ks.test(data, "pnorm", mean = mean(data), sd = sd(data))
    
    # Perform Anderson-Darling test
    ad_result <- ad.test(data)
    
    # Append the results to the dataframe
    normality_results <- rbind(normality_results, data.frame(
      Dataset = dataset_name,
      Variable = var,
      SW_p = formatC(shapiro_result$p.value, format = "f", digits = 5),
      KS_p = formatC(ks_result$p.value, format = "f", digits = 5),
      AD_p = formatC(ad_result$p.value, format = "f", digits = 5)
    ))
  }
}

# Order the results by Variable
normality_results <- normality_results[order(normality_results$Variable), ]

# Display the results
print(normality_results)
```

### Histograms of Response Variables

```{r}
#| label: Histograms of 10 variables of interest for 1 year comparison data
#| warning: false

# Specify the variables
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", 
              "totalshots", "totaloff",  
               "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create a list to store the plots
plots <- list()

# Generate a histogram for each variable and store it in the list
for (var in variables) {
  p <- ggplot(EPL_1, aes_string(x = var)) + 
    geom_histogram(binwidth = 1, fill = "blue", color = "black") +
    theme_minimal()
  plots[[var]] <- p
}

# Arrange the plots in a 4x3 grid
do.call("grid.arrange", c(plots, ncol = 3))
```

```{r}
#| label: Histograms of 10 variables of interest for 3 year comparison data
#| warning: false

# Specify the variables
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", 
               "totalshots", "totaloff",  
               "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create a list to store the plots
plots <- list()

# Generate a histogram for each variable and store it in the list
for (var in variables) {
  p <- ggplot(EPL_3, aes_string(x = var)) + 
    geom_histogram(binwidth = 1, fill = "blue", color = "black") +
    theme_minimal()
  plots[[var]] <- p
}

# Arrange the plots in a 4x3 grid
do.call("grid.arrange", c(plots, ncol = 3))
```

```{r}
#| label: Histograms of 10 variables of interest for 5 year comparison data
#| warning: false

# Specify the variables
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", 
               "totalshots", "totaloff",  
               "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create a list to store the plots
plots <- list()

# Generate a histogram for each variable and store it in the list
for (var in variables) {
  p <- ggplot(EPL_5, aes_string(x = var)) + 
    geom_histogram(binwidth = 1, fill = "blue", color = "black") +
    theme_minimal()
  plots[[var]] <- p
}

# Arrange the plots in a 4x3 grid
do.call("grid.arrange", c(plots, ncol = 3))
```

## Checking Equal Variance Assumption

### BP and Levene's Tests

```{r}
#| label: Brusch-Pagan and Levene`s Tests for Normality
#| warning: false


# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create an empty dataframe to store the results
eqvar_results <- data.frame(Variable = character(), levTest_1p = numeric(), levTest_3p = numeric(), levTest_5p = numeric(), bpTest_1p = numeric(), bpTest_3p = numeric(), bpTest_5p = numeric(), stringsAsFactors = FALSE)

# Loop over each variable
for (var in variables) {
  
  # Levene Test for EPL_1
  levTest1 <- leveneTest(get(var) ~ as.factor(PostVAR), data = EPL_1)
  levP1 <- levTest1$`Pr(>F)`[1]
  
  # Levene Test for EPL_3
  levTest3 <- leveneTest(get(var) ~ as.factor(PostVAR), data = EPL_3)
  levP3 <- levTest3$`Pr(>F)`[1]
  
  # Levene Test for EPL_5
  levTest5 <- leveneTest(get(var) ~ as.factor(PostVAR), data = EPL_5)
  levP5 <- levTest5$`Pr(>F)`[1]

  # Breusch-Pagan Test for EPL_1
  bpTest1 <- bptest(get(var) ~ PostVAR, data = EPL_1)
  bpP1 <- bpTest1$p.value

  # Breusch-Pagan Test for EPL_3
  bpTest3 <- bptest(get(var) ~ PostVAR, data = EPL_3)
  bpP3 <- bpTest3$p.value

  # Breusch-Pagan Test for EPL_5
  bpTest5 <- bptest(get(var) ~ PostVAR, data = EPL_5)
  bpP5 <- bpTest5$p.value
  
  # Append the results to the dataframe
  eqvar_results <- rbind(eqvar_results, data.frame(
    Variable = var,
    levTest_1p = formatC(levP1, format = "f", digits = 5),
    levTest_3p = formatC(levP3, format = "f", digits = 5),
    levTest_5p = formatC(levP5, format = "f", digits = 5),
    bpTest_1p = formatC(bpP1, format = "f", digits = 5),
    bpTest_3p = formatC(bpP3, format = "f", digits = 5),
    bpTest_5p = formatC(bpP5, format = "f", digits = 5)
  ))
}

# Display the results
print(eqvar_results)
```

## -

## Welch's T-Tests

```{r}
#| label: T-Test Analysis
#| warning: false

# Prevent scientific notation for p-values
options(scipen = 999)

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create an empty list to store t-test results
t_test_results <- list()

# Loop over each variable
for (var in variables) {
  # 1 year
  t_test_1 <- t.test(get(var) ~ PostVAR, var.equal = FALSE, data = EPL_1)
  # 3 years
  t_test_3 <- t.test(get(var) ~ PostVAR, var.equal = FALSE, data = EPL_3)
  # 5 years
  t_test_5 <- t.test(get(var) ~ PostVAR, var.equal = FALSE, data = EPL_5)
  
  # Extract and tidy t-test results using broom
  tidy_1 <- tidy(t_test_1)
  tidy_1$Years <- "1 year"
  tidy_3 <- tidy(t_test_3)
  tidy_3$Years <- "3 years"
  tidy_5 <- tidy(t_test_5)
  tidy_5$Years <- "5 years"
  
  # Combine results into one dataframe
  result_df <- bind_rows(tidy_1, tidy_3, tidy_5) %>%
    mutate(Variable = var) %>%
    relocate(Variable, Years)
  
  # Store results in a list
  t_test_results[[var]] <- result_df
}

# Combine all results into one dataframe
t_test_result <- bind_rows(t_test_results)

# Round all numeric columns to 8 decimal places
t_test_result <- t_test_result %>%
  mutate(across(where(is.numeric), round, 8))

# View the final results
print(t_test_result)
```

## Mann-Whitney U-Tests

```{r}
#| label: Mann-Whitney U-Test Analysis
#| warning: false

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create an empty dataframe to store all t-test results
MW_U_results <- data.frame()

# Loop over each variable
for (var in variables) {
  # 1 year
  t_test_1 <- wmwTest(get(var) ~ PostVAR, data = EPL_1)
  # 3 years
  t_test_3 <- wmwTest(get(var) ~ PostVAR, data = EPL_3)
  # 5 years
  t_test_5 <- wmwTest(get(var) ~ PostVAR, data = EPL_5)
  
  # Extract relevant information and store in data frame
  result_df <- data.frame(
    Variable = var,
    Test = c("Estimate", "Conf.Lower", "Conf.Upper", "P.Value"),
    `1_year` = c(t_test_1$estimate, t_test_1$conf.int[1], t_test_1$conf.int[2], t_test_1$p.value),
    `3_years` = c(t_test_3$estimate, t_test_3$conf.int[1], t_test_3$conf.int[2], t_test_3$p.value),
    `5_years` = c(t_test_5$estimate, t_test_5$conf.int[1], t_test_5$conf.int[2], t_test_5$p.value)
  )
  
  # Append results to the master dataframe
  MW_U_results <- bind_rows(MW_U_results, result_df[4,])
}
```

## Generalized Linear Models

```{r}
#| label: Generalized Linear Models analysis
#| warning: false

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Create an empty dataframe to store all t-test results
GLM_results <- data.frame()

# Loop over each variable
for (var in variables) {
  # 1 year
  glm_1 <- glm.nb(get(var) ~ PostVAR, link = "log", data = EPL_1)
  # 3 years
  glm_3 <- glm.nb(get(var) ~ PostVAR, link = "log", data = EPL_3)
  # 5 years
  glm_5 <- glm.nb(get(var) ~ PostVAR, link = "log", data = EPL_5)
  
  # Extract relevant information and store in data frame
  result_df <- data.frame(
    Variable = var,
    Test = c("p-value", "AIC"),
    `1_year` = c(summary(glm_1)$coefficients[8], summary(glm_1)$aic),
    `3_years` = c(summary(glm_3)$coefficients[8], summary(glm_3)$aic),
    `5_years` = c(summary(glm_5)$coefficients[8], summary(glm_5)$aic)
  )
  
  # Append results to the master dataframe
  GLM_results <- bind_rows(GLM_results, result_df[1:2,])
}

print(GLM_results)
```

```{r}
#| label: Generalized Linear Models Confidence Intervals

# List of variables to perform GLM and compute confidence intervals
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", 
                "totalshots", "totaloff", "totalred", 
               "totalyellow", "totaltackleswon", "totalint")

# List of datasets
datasets <- list(EPL_1 = EPL_1, EPL_3 = EPL_3, EPL_5 = EPL_5)

# Initialize a list to store all results
all_results <- list()

# Outer loop over each dataset
for (dataset_name in names(datasets)) {
  dataset <- datasets[[dataset_name]]
  
  # Initialize a list to store results for the current dataset
  dataset_results <- list()
  
  # Inner loop over each variable
  for (var in variables) {
    # Fit the GLM with negative binomial distribution
    glm_model <- glm.nb(as.formula(paste(var, "~ as.factor(PostVAR)")), data = dataset, link = "log")
    
    # Predict the mean response for PostVAR = 0 and PostVAR = 1 on the link scale
    new_data <- data.frame(PostVAR = c(0, 1))
    predictions <- predict(glm_model, newdata = new_data, type = "link", se.fit = TRUE)
    
    # Extract the fitted values and their standard errors on the link scale
    fit_values <- predictions$fit
    se_values <- predictions$se.fit
    
    # Compute the critical value for the 95% confidence interval
    alpha <- 0.05
    z <- qnorm(1 - alpha / 2)
    
    # Calculate the confidence intervals on the link scale
    link_lower <- fit_values - z * se_values
    link_upper <- fit_values + z * se_values
    
    # Transform the predictions and confidence intervals back to the response scale
    response_fit <- glm_model$family$linkinv(fit_values)
    response_lower <- glm_model$family$linkinv(link_lower)
    response_upper <- glm_model$family$linkinv(link_upper)
    
    # Calculate the difference and its standard error
    diff_fit <- response_fit[2] - response_fit[1]
    diff_se <- sqrt((response_fit[2] * se_values[2])^2 + (response_fit[1] * se_values[1])^2)
    
    # Compute the confidence interval for the difference
    diff_lower <- diff_fit - z * diff_se
    diff_upper <- diff_fit + z * diff_se
    
    # Combine results into a data frame
    results <- data.frame(
      Dataset = dataset_name,
      Variable = var,
      PostVAR = c(0, 1),
      fitted_values = response_fit,
      lower_bound = response_lower,
      upper_bound = response_upper
    )
    
    # Add the difference and its confidence interval
    GLM_CIs <- data.frame(
      Dataset = dataset_name,
      Variable = var,
      difference = diff_fit,
      lower_bound = diff_lower,
      upper_bound = diff_upper
    )
    
    # Store the results in the list for the current dataset
    dataset_results[[var]] <- list(predictions = results, difference = GLM_CIs)
  }
  
  # Store the results for the current dataset in the overall results list
  all_results[[dataset_name]] <- dataset_results
}

# Combine all results into a single data frame for predictions and differences
GLM_predictions <- do.call(rbind, lapply(all_results, function(ds) do.call(rbind, lapply(ds, function(x) x$predictions))))
GLM_CIs <- do.call(rbind, lapply(all_results, function(ds) do.call(rbind, lapply(ds, function(x) x$difference))))

# Display the results
print(GLM_predictions)
print(GLM_CIs)
```

## -

## Appendix

#### Separate Analysis for Corner Kicks

```{r}
#| label: Separate Analysis for Corner Kicks

#2 year comparison dataset (FBref didn't supply number of corners until 17/18 season)
EPL_2 <- EPL_Totals %>%
  filter(Season %in% c("17/18", "18/19", "19/20", "20/21"))

#T-Test analysis
t.test(totalcorners ~ PostVAR, eq.var = FALSE, data = EPL_2)

#Mann-Whitney U-Test analysis
wmwTest(totalcorners ~ PostVAR, data = EPL_2)

#Generalized Linear Model analysis
summary(glm.nb(totalcorners ~ PostVAR, link = "log", data = EPL_2))$coefficients[8]
AIC(glm.nb(totalcorners ~ PostVAR, link = "log", data = EPL_2))
```

#### Z-Tests (not used in paper)

```{r}
#| label: Z-Test analysis, using SD values from EPL_Totals (EPL_5) dataset

z_test <- function(sample_data, post_var, population_sd) {
  # Split data into two groups based on PostVAR
  group1 <- sample_data[post_var == 0]
  group2 <- sample_data[post_var == 1]
  
  # Calculate means and sample sizes
  mean1 <- mean(group1, na.rm = TRUE)
  mean2 <- mean(group2, na.rm = TRUE)
  n1 <- length(group1)
  n2 <- length(group2)
  
  # Calculate z-statistic
  z_statistic <- (mean2 - mean1) / sqrt((population_sd^2 / n1) + (population_sd^2 / n2))
  
  # Calculate p-value
  p_value <- 2 * (1 - pnorm(abs(z_statistic)))
  
  # Confidence interval (assuming a normal distribution)
  conf_int <- (mean2 - mean1) + c(-1, 1) * qnorm(0.975) * sqrt((population_sd^2 / n1) + (population_sd^2 / n2))
  
  list(estimate = mean2 - mean1, conf.int = conf_int, p.value = p_value)
}

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")


# Assume a population standard deviation for each variable
sd_pop <- c(
  totalgoals = 1.6434029, totalfoulscom = 5.1740743, totalpksatt = 0.5122013, totalpksmade = 0.4553075, 
  totalshots = 5.7816171, totaloff = 2.2169412, totalred = 0.3562214, totalyellow = 2.0222872, totaltackleswon = 7.9990457, totalint = 8.3870906)

# Create an empty list to store z-test results
z_tests <- list()

# Loop over each variable
for (var in variables) {
  # 1 year
  z_test_1 <- z_test(EPL_1[[var]], EPL_1$PostVAR, sd_pop[[var]])
  # 3 years
  z_test_3 <- z_test(EPL_3[[var]], EPL_3$PostVAR, sd_pop[[var]])
  # 5 years
  z_test_5 <- z_test(EPL_5[[var]], EPL_5$PostVAR, sd_pop[[var]])
  
  # Extract z-test estimate, 95% CI, and p-value, and store in data frame
  result_df <- data.frame(
    Variable = var,
    Statistic = c("Estimate", "Lower CI limit", "Upper CI limit", "P-value"),
    `1 year` = format(c(z_test_1$estimate, z_test_1$conf.int[1], z_test_1$conf.int[2], z_test_1$p.value), scientific = FALSE, digits = 5),
    `3 years` = format(c(z_test_3$estimate, z_test_3$conf.int[1], z_test_3$conf.int[2], z_test_3$p.value), scientific = FALSE, digits = 5),
    `5 years` = format(c(z_test_5$estimate, z_test_5$conf.int[1], z_test_5$conf.int[2], z_test_5$p.value), scientific = FALSE, digits = 5)
  )
  
  # Append the result dataframe to the list
  z_tests[[var]] <- result_df
}

# Combine all result data frames into one
z_test_results <- do.call(rbind, z_tests)
```

```{r}
#| label: Z-Test analysis, using SD values from respective datasets

z_test <- function(sample_data, post_var, population_sd) {
  # Split data into two groups based on PostVAR
  group1 <- sample_data[post_var == 0]
  group2 <- sample_data[post_var == 1]
  
  # Calculate means and sample sizes
  mean1 <- mean(group1, na.rm = TRUE)
  mean2 <- mean(group2, na.rm = TRUE)
  n1 <- length(group1)
  n2 <- length(group2)
  
  # Calculate z-statistic
  z_statistic <- (mean2 - mean1) / sqrt((population_sd^2 / n1) + (population_sd^2 / n2))
  
  # Calculate p-value
  p_value <- 2 * (1 - pnorm(abs(z_statistic)))
  
  # Confidence interval (assuming a normal distribution)
  conf_int <- (mean2 - mean1) + c(-1, 1) * qnorm(0.975) * sqrt((population_sd^2 / n1) + (population_sd^2 / n2))
  
  list(estimate = mean2 - mean1, conf.int = conf_int, p.value = p_value)
}

# List of variables of interest
variables <- c("totalgoals", "totalfoulscom", "totalpksatt", "totalpksmade", "totalshots", "totaloff", "totalred", "totalyellow", "totaltackleswon", "totalint")

# Assume a population standard deviation for each variable for EPL_1
sd_1 <- c(
  totalgoals = 1.5380413, totalfoulscom = 5.3360648, totalpksatt = 0.5016010, totalpksmade = 0.4562246, 
  totalshots = 5.5532647, totaloff = 2.2630061, totalred = 0.3459976, totalyellow = 1.9380103, totaltackleswon = 7.7276553, totalint = 5.4362745)

# Assume a population standard deviation for each variable for EPL_33
sd_3 <- c(
  totalgoals = 1.6219835, totalfoulscom = 5.2691893, totalpksatt = 0.5138992, totalpksmade = 0.4578733, 
  totalshots = 5.5799063, totaloff = 2.2483760, totalred = 0.3407801, totalyellow = 1.9827667, totaltackleswon = 7.8532875, totalint = 6.6577616)

# Assume a population standard deviation for each variable for EPL_5
sd_5 <- c(
  totalgoals = 1.6434029, totalfoulscom = 5.1740743, totalpksatt = 0.5122013, totalpksmade = 0.4553075,
  totalshots = 5.7816171, totaloff = 2.2169412, totalred = 0.3562214, totalyellow = 2.0222872, totaltackleswon = 7.9990457, totalint = 8.3870906)

# Create an empty list to store z-test results
z_tests <- list()

# Loop over each variable
for (var in variables) {
  # 1 year
  z_test_1 <- z_test(EPL_1[[var]], EPL_1$PostVAR, sd_1[[var]])
  # 3 years
  z_test_3 <- z_test(EPL_3[[var]], EPL_3$PostVAR, sd_3[[var]])
  # 5 years
  z_test_5 <- z_test(EPL_5[[var]], EPL_5$PostVAR, sd_5[[var]])
  
  # Extract z-test estimate, 95% CI, and p-value, and store in data frame
  result_df <- data.frame(
    Variable = var,
    Statistic = c("Estimate", "Lower CI limit", "Upper CI limit", "P-value"),
    `1 year` = format(c(z_test_1$estimate, z_test_1$conf.int[1], z_test_1$conf.int[2], z_test_1$p.value), scientific = FALSE, digits = 5),
    `3 years` = format(c(z_test_3$estimate, z_test_3$conf.int[1], z_test_3$conf.int[2], z_test_3$p.value), scientific = FALSE, digits = 5),
    `5 years` = format(c(z_test_5$estimate, z_test_5$conf.int[1], z_test_5$conf.int[2], z_test_5$p.value), scientific = FALSE, digits = 5)
  )
  
  # Append the result dataframe to the list
  z_tests[[var]] <- result_df
}

# Combine all result data frames into one
z_test_results <- do.call(rbind, z_tests)
```
